---
output:
  github_document:
    toc: false
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Edit README.Rmd, not README.md. -->
```{r setup, include=FALSE}
pkgload::load_all(
  path = ".",
  export_all = FALSE,
  helpers = FALSE,
  quiet = TRUE
)
```


# gp3bayes

`gp3bayes` is an independent R package under development for transparent,
contract-first Bayesian workflows for repeated-measures and hierarchical
behavioural data.

## Scope

The package is designed around:

- explicit data and model contracts;
- model-readiness audits;
- deterministic simulation with recorded generating parameters;
- inspectable prior specifications;
- prior and posterior predictive checks;
- convergence and sampling diagnostics;
- prior-sensitivity assessment;
- parameter-recovery workflows;
- conservative and reproducible reporting.

The initial development scope is restricted to:

1. hierarchical Bernoulli-logit models for binary trial-level outcomes;
2. hierarchical lognormal models for strictly positive uncensored durations.

The core package will not require Gazepoint hardware, Gazepoint exports,
`gp3tools`, proprietary software, private data, or a Bayesian backend.

## Model contracts

`create_model_contract()` records the approved methodological specification
and neutral column mappings for one initial model family. Creating a
contract does not validate data, fit a model, or establish model adequacy.

```{r model-contract-example}
binary_contract <- create_model_contract(
  family = "binary",
  outcome_col = "selected",
  participant_col = "participant_id",
  item_col = "stimulus_id",
  trial_col = "trial_id",
  condition_col = "condition"
)

binary_contract
```

## Readiness audits

`audit_model_readiness()` evaluates observable data requirements before
formula construction or model fitting. Failures block progression, whereas
warnings identify structures requiring review.

```{r readiness-audit-example}
binary_data <- data.frame(
  participant_id = rep(c("p1", "p2"), each = 4),
  stimulus_id = rep(paste0("s", 1:4), times = 2),
  trial_id = rep(1:4, times = 2),
  condition = rep(c("control", "treatment"), times = 4),
  selected = c(0, 1, 0, 1, 1, 0, 1, 0)
)

readiness_audit <- audit_model_readiness(
  binary_data,
  binary_contract
)

readiness_audit
```

## Model specifications

`build_model_formula()` translates the approved contract into an R formula,
while `create_prior_specification()` records family-appropriate priors
without creating backend-specific objects. A ready audit, formula, contract,
and validated priors can then be combined into one inspectable model
specification.

```{r model-specification-example}
binary_priors <- create_prior_specification(
  binary_contract,
  baseline = 0.5
)

binary_specification <- create_model_specification(
  binary_contract,
  readiness_audit,
  binary_priors
)

binary_specification
```

## Development status

`gp3bayes` is currently at development version `0.0.0.9000`.

No public model-fitting API is available yet. The current development stage
establishes the standalone package structure, approved methodological scope,
data-contract principles, and validation infrastructure.

## Interpretation boundaries

Behavioural, gaze, pupil, and physiological measurements do not directly
reveal emotion, stress, cognition, comprehension, personality, diagnosis,
deception, intention, or other latent psychological states.

Associations must not be described as causal effects unless the study design
and target estimand justify causal interpretation.

## Licence

`gp3bayes` is released under the MIT License.
