---
title: "Bayesian Multilevel Gaze Mediation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bayesian Multilevel Gaze Mediation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Model contract

`gp3bayes` consumes a canonical preparation object from `eyeprocess`. It does not recreate within/between components.

```{r, eval=FALSE}
priors <- create_mediation_prior_specification(
  coefficient_sd = 0.75,
  group_sd_scale = 1
)

spec <- specify_multilevel_gaze_mediation(
  prepared,
  mediator_family = "gaussian",
  outcome_family = "bernoulli",
  priors = priors,
  random_slopes = "mediator_x",
  missingness_policy = "error"
)
```

Paths without observed design variation are not manufactured. In a perfectly balanced within-subject manipulation, the between-person X component may be constant and the corresponding between-X paths are reported as not estimable.

## Fit and diagnose

```{r, eval=FALSE}
fit <- fit_multilevel_gaze_mediation(
  prepared,
  mediator_family = "gaussian",
  outcome_family = "bernoulli",
  priors = priors,
  random_slopes = "mediator_x",
  missingness_policy = "error"
)
check_mediation_convergence(fit)
estimate_within_indirect_effect(fit)
posterior_predictive_check_mediation(fit)
```

Indirect effects are blocked by default after critical convergence failure. For nonlinear families the coefficient product is on the linear-predictor product scale, not a probability-scale natural indirect effect.

## Extensions

Serial and moderated mediation require components prepared upstream in `eyeprocess`. Do not fit arbitrary chains without defensible temporal/design ordering.


## Prior sensitivity

Do not treat the default priors as invisible. Refit the same design under a defensible narrower and wider coefficient prior, then compare the indirect-effect distribution, posterior predictive behavior, and convergence diagnostics.

```{r, eval=FALSE}
narrow <- create_mediation_prior_specification(coefficient_sd = 0.5)
wide <- create_mediation_prior_specification(coefficient_sd = 1.5)

fit_narrow <- fit_multilevel_gaze_mediation(
  prepared, priors = narrow, random_slopes = "mediator_x"
)
fit_wide <- fit_multilevel_gaze_mediation(
  prepared, priors = wide, random_slopes = "mediator_x"
)
```

Serial and moderated extensions currently use participant random intercepts; requests for extension-specific random slopes are rejected rather than silently ignored.
