---
title: "Bayesian Multilevel Gaze Mediation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bayesian Multilevel Gaze Mediation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Model contract

`gp3bayes` consumes a canonical preparation object from `eyeprocess`. It does not recreate within/between components.

```{r, eval=FALSE}
priors <- create_mediation_prior_specification(
  coefficient_sd = 0.75,
  group_sd_scale = 1
)

spec <- specify_multilevel_gaze_mediation(
  prepared,
  mediator_family = "gaussian",
  outcome_family = "bernoulli",
  priors = priors,
  random_slopes = "mediator_x",
  missingness_policy = "error"
)
```

Paths without observed design variation are not manufactured. In a perfectly balanced within-subject manipulation, the between-person X component may be constant and the corresponding between-X paths are reported as not estimable.

## Fit and diagnose

```{r, eval=FALSE}
fit <- fit_multilevel_gaze_mediation(
  prepared,
  mediator_family = "gaussian",
  outcome_family = "bernoulli",
  priors = priors,
  random_slopes = "mediator_x",
  missingness_policy = "error"
)
check_mediation_convergence(fit)
estimate_within_indirect_effect(fit)
posterior_predictive_check_mediation(fit)
```

Indirect effects are blocked by default after critical convergence failure. For nonlinear families the coefficient product is on the linear-predictor product scale, not a probability-scale natural indirect effect.

## Extensions

Serial and moderated mediation require components prepared upstream in `eyeprocess`. Do not fit arbitrary chains without defensible temporal/design ordering.


## Prior sensitivity

Do not treat the default priors as invisible. Refit the same design under a defensible narrower and wider coefficient prior, then compare the indirect-effect distribution, posterior predictive behavior, and convergence diagnostics.

```{r, eval=FALSE}
narrow <- create_mediation_prior_specification(coefficient_sd = 0.5)
wide <- create_mediation_prior_specification(coefficient_sd = 1.5)

fit_narrow <- fit_multilevel_gaze_mediation(
  prepared, priors = narrow, random_slopes = "mediator_x"
)
fit_wide <- fit_multilevel_gaze_mediation(
  prepared, priors = wide, random_slopes = "mediator_x"
)
```

Serial and moderated extensions currently use participant random intercepts; requests for extension-specific random slopes are rejected rather than silently ignored.


## Assumptions and limitations

Separating within- and between-participant paths prevents level conflation, but a fitted coefficient product does not automatically identify a causal natural indirect effect. Causal language requires a defensible temporal order and assumptions about exposure–mediator and mediator–outcome confounding. For Bernoulli, ordinal, and count outcomes, the implemented coefficient-product indirect effect is on the model's linear-predictor product scale.

Missingness policies are explicit but do not solve informative missingness. If gaze loss differs by condition, participant, stimulus, or outcome tendency, report those patterns and examine defensible sensitivity specifications. A passed convergence check also does not prove model adequacy; posterior predictive checks and prior sensitivity remain necessary.

## Reporting example

A concise report should state the trial-level observational unit, within/between decomposition, mediator and outcome families, random-effects structure, missingness and quality policy, priors, sampler settings, convergence diagnostics, posterior predictive checks, and the scale of the indirect effect.

For a binary outcome, suitable wording is:

> The within-participant indirect effect (`a_W * b_W`) was summarized on the linear-predictor product scale. Missing gaze was not coded as zero. Between-person exposure paths were reported only when the observed design contained between-person exposure variation. The indirect-effect summary was extracted only after the declared convergence gate passed.

Use `report_multilevel_gaze_mediation()` as a structured starting point, then add design-specific interpretation, sensitivity results, and limitations. Do not convert a logit-scale coefficient product into a percentage-point mediated effect without a separately implemented predictive/counterfactual estimand.

## API links

The core workflow uses `create_mediation_prior_specification()`, `specify_multilevel_gaze_mediation()`, `fit_multilevel_gaze_mediation()`, `check_mediation_convergence()`, `posterior_predictive_check_mediation()`, `estimate_within_indirect_effect()`, `estimate_between_indirect_effect()`, `plot_indirect_effect_distribution()`, and `report_multilevel_gaze_mediation()`. Serial and moderated extensions are available through their explicitly named specification/fitting functions when the design supports those structures.
