---
title: "End-to-End Hierarchical Binary Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{End-to-End Hierarchical Binary Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Scope

This article presents the approved Bernoulli-logit workflow for binary
trial-level outcomes with repeated observations, participant effects, and
optional crossed item effects. The interface is deliberately restricted:
users do not supply an arbitrary formula, likelihood, backend, Stan program,
or automatic model-selection rule.

The core contract, simulation, preparation, model specification, and prior
predictive check do not require a Bayesian backend.

## Simulate a known data-generating process

```{r binary-simulation}
library(gp3bayes)

binary_simulation <- simulate_hierarchical_binary_data(
  n_participants = 16,
  trials_per_participant = 10,
  n_items = 8,
  random_slope_sd = 0.20,
  seed = 2026
)

binary_simulation
head(binary_simulation$data)
```

The simulation object retains fixed effects, grouping scales, the condition
coding, random effects, and the random-number seed. Stored truth enables
later recovery assessment without a private or external data set.

## Declare the model contract

```{r binary-contract}
binary_contract <- create_model_contract(
  family = "binary",
  outcome_col = "selected",
  participant_col = "participant_id",
  item_col = "item_id",
  trial_col = "trial_id",
  condition_col = "condition",
  predictors = c(
    "participant_covariate",
    "trial_covariate"
  ),
  interaction = c(
    "condition",
    "participant_covariate"
  ),
  random_slope = TRUE
)

binary_contract
```

Contract creation records the intended outcome, grouping structure,
likelihood, link, supported interaction, assumptions, diagnostics, limitations,
and interpretation boundaries. It does not validate the data or establish
model adequacy.

## Prepare and audit the data

```{r binary-preparation}
binary_prepared <- prepare_hierarchical_binary_data(
  binary_simulation$data,
  binary_contract,
  condition_levels = c(
    "control",
    "treatment"
  )
)

binary_prepared
binary_prepared$decision_log
```

Outcome mapping, condition coding, missing-value decisions, and requested
scaling are explicit and recorded. No variable is silently transformed.

## Specify priors and inspect prior implications

```{r binary-specification}
binary_specification <- specify_binary_model(
  binary_prepared,
  baseline = 0.35,
  intercept_scale = 1.5,
  coefficient_scale = 0.75,
  group_sd_scale = 1,
  correlation_eta = 2,
  student_df = 3
)

binary_specification
binary_specification$priors$table
```

```{r binary-prior-predictive}
binary_prior_predictive <- check_binary_prior_predictive(
  binary_specification,
  draws = 100,
  seed = 2027
)

binary_prior_predictive
binary_prior_predictive$checks
```

A prior-predictive failure requests substantive review. The function does not
automatically change priors or select a different model.

## Translate and fit with the optional backend

The fitting route is fixed to `brms`, `rstan`, and full MCMC sampling. The
following code requires the optional backend and a working C++ toolchain.

```{r binary-fit, eval=FALSE}
binary_translation <- translate_binary_model_to_brms(
  binary_specification
)

binary_fit <- fit_binary_model(
  binary_specification,
  chains = 4,
  iter = 2000,
  warmup = 1000,
  cores = 4,
  seed = 2028,
  adapt_delta = 0.95,
  max_treedepth = 12,
  refresh = 100
)
```

A returned fit confirms that sampling completed. It does not by itself
establish convergence or posterior adequacy.

## Diagnose, interpret, and validate

```{r binary-validation, eval=FALSE}
binary_diagnostics <- diagnose_binary_fit(
  binary_fit
)

binary_posterior <- summarise_binary_posterior(
  binary_fit,
  probability = 0.95
)

binary_predictive <- check_binary_posterior_predictive(
  binary_fit,
  draws = 500,
  seed = 2029
)

binary_diagnostics
binary_posterior
binary_predictive
```

The diagnostic object reports R-hat, bulk and tail ESS, divergent transitions,
maximum-treedepth saturation, and chain-level energy diagnostics. Its
pass/review/fail status is a threshold report, not an automatic declaration
that the model converged.

Population-level coefficients are reported on the log-odds and odds-ratio
scales. Posterior probabilities and intervals are not frequentist significance
tests and are not automatically causal.

## Prior sensitivity and simulation recovery

```{r binary-sensitivity-recovery, eval=FALSE}
binary_sensitivity <- assess_binary_prior_sensitivity(
  binary_fit,
  scale_multipliers = c(
    tighter = 0.5,
    wider = 2
  )
)

binary_recovery <- run_binary_recovery(
  repetitions = 20,
  seed = 3001
)
```

A small recovery run is only a smoke test. A larger run assesses the declared
synthetic design and does not validate every future use of the package.

## Structured reporting

```{r binary-report, eval=FALSE}
binary_report <- create_binary_model_report(
  binary_fit,
  diagnostics = binary_diagnostics,
  posterior_summary = binary_posterior,
  posterior_predictive = binary_predictive,
  prior_sensitivity = binary_sensitivity,
  recovery = binary_recovery,
  file = "binary-model-report.md"
)

binary_report
```

The report keeps fitting, diagnostics, predictive checks, sensitivity, and
recovery as separate evidence layers. It does not automatically claim
convergence, predictive validity, causal identification, or substantive
validity.
