---
title: "End-to-End Hierarchical Lognormal Duration Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{End-to-End Hierarchical Lognormal Duration Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Scope

The duration workflow is restricted to strictly positive, finite, uncensored
durations modeled with a hierarchical lognormal likelihood. Zero, negative,
censored, truncated, shifted-lognormal, Gamma, Weibull, survival, and mixture
outcomes are outside this contract.

## Simulate durations with stored truth

```{r duration-simulation}
library(gp3bayes)

duration_simulation <- simulate_hierarchical_duration_data(
  n_participants = 16,
  trials_per_participant = 10,
  n_items = 8,
  baseline_median = 500,
  random_slope_sd = 0.15,
  residual_sd = 0.40,
  outcome_unit = "milliseconds",
  seed = 2030
)

duration_simulation
head(duration_simulation$data)
```

The fixed effects and grouping scales are on the log-duration scale. The
baseline is stored as a median in the declared outcome unit.

## Declare the duration contract

```{r duration-contract}
duration_contract <- create_model_contract(
  family = "duration",
  outcome_col = "duration",
  participant_col = "participant_id",
  item_col = "item_id",
  trial_col = "trial_id",
  condition_col = "condition",
  predictors = c(
    "participant_covariate",
    "trial_covariate"
  ),
  interaction = c(
    "condition",
    "participant_covariate"
  ),
  random_slope = TRUE,
  outcome_unit = "milliseconds"
)

duration_contract
```

The outcome unit is mandatory. The package never guesses a unit from column
names or magnitudes.

## Prepare, convert, and audit

```{r duration-preparation}
duration_prepared <- prepare_hierarchical_duration_data(
  duration_simulation$data,
  duration_contract,
  condition_levels = c(
    "control",
    "treatment"
  )
)

duration_prepared
duration_prepared$decision_log
```

Explicit unit conversion uses `outcome_multiplier` and `converted_unit`
together. For example, converting milliseconds to seconds is recorded as:

```{r duration-unit-conversion}
duration_seconds <- prepare_hierarchical_duration_data(
  duration_simulation$data,
  duration_contract,
  condition_levels = c(
    "control",
    "treatment"
  ),
  outcome_multiplier = 0.001,
  converted_unit = "seconds"
)

duration_seconds$transformations$outcome
```

## Specify priors and check prior implications

```{r duration-specification}
duration_specification <- specify_duration_model(
  duration_prepared,
  baseline = 500,
  intercept_scale = 1,
  coefficient_scale = 0.5,
  group_sd_scale = 1,
  residual_scale = 1,
  correlation_eta = 2,
  student_df = 3
)

duration_specification
duration_specification$priors$table
```

```{r duration-prior-predictive}
duration_prior_predictive <- check_duration_prior_predictive(
  duration_specification,
  draws = 100,
  seed = 2031
)

duration_prior_predictive
duration_prior_predictive$checks
```

The prior check examines overall medians, upper tails, coefficients of
variation, and condition median ratios. A failure requests review and does not
automatically alter the priors.

## Translate and fit through the restricted backend

```{r duration-fit, eval=FALSE}
duration_translation <- translate_duration_model_to_brms(
  duration_specification
)

duration_fit <- fit_duration_model(
  duration_specification,
  chains = 4,
  iter = 2000,
  warmup = 1000,
  cores = 4,
  seed = 2032,
  adapt_delta = 0.95,
  max_treedepth = 12,
  refresh = 100
)
```

The likelihood, formula, priors, backend, and sampling algorithm are derived
from the approved package specification. There is no unrestricted formula or
backend argument.

## Diagnostics and posterior interpretation

```{r duration-validation, eval=FALSE}
duration_diagnostics <- diagnose_duration_fit(
  duration_fit
)

duration_posterior <- summarise_duration_posterior(
  duration_fit
)

duration_predictive <- check_duration_posterior_predictive(
  duration_fit,
  draws = 500,
  seed = 2033
)
```

Exponentiating a population-level coefficient gives a conditional median
duration ratio under the lognormal model. This ratio is not automatically a
causal effect.

The predictive check compares observed and replicated median, mean, upper-tail,
dispersion, condition-ratio, and grouping summaries. Passing those summaries
does not prove global model adequacy.

## Sensitivity, recovery, and reporting

```{r duration-sensitivity-recovery, eval=FALSE}
duration_sensitivity <- assess_duration_prior_sensitivity(
  duration_fit,
  scale_multipliers = c(
    tighter = 0.5,
    wider = 2
  )
)

duration_recovery <- run_duration_recovery(
  repetitions = 20,
  baseline_median = 500,
  outcome_unit = "milliseconds",
  seed = 4001
)

duration_report <- create_duration_model_report(
  duration_fit,
  diagnostics = duration_diagnostics,
  posterior_summary = duration_posterior,
  posterior_predictive = duration_predictive,
  prior_sensitivity = duration_sensitivity,
  recovery = duration_recovery,
  file = "duration-model-report.md"
)
```

Recovery results apply to the declared synthetic data-generating process.
Reports preserve the distinction between successful fitting, numerical
sampling diagnostics, predictive behavior, robustness checks, and substantive
interpretation.
