---
title: "Mediation evidence atlas: effect, bootstrap, DAG figures"
output: rmarkdown::html_vignette
vignette: >
  %\\VignetteIndexEntry{Mediation evidence atlas}
  %\\VignetteEngine{knitr::rmarkdown}
  %\\VignetteEncoding{UTF-8}
---

## Why these figures

The open-source [BriDGE research package](https://github.com/gav888/BriDGE) inspired the **types** of figure adapted here: effect-size comparison, bootstrap distributions and a researcher-versus-discovered causal-edge audit. The implementation is original, and it does not replicate the original mediation estimator or offer automatic causal discovery.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

### Supplied effect estimates

```{r effects, eval=FALSE}
library(gp3bayes)
est <- data.frame(
  term = c("Gaze indirect", "Pupil indirect"),
  estimate = c(0.12, -0.05),
  lower = c(0.03, -0.14),
  upper = c(0.22, 0.04)
)
plot_mediation_evidence_atlas(est, type = "effects")
```

### Bootstrap or posterior draws

```{r draws, eval=FALSE}
set.seed(2026)
draws <- data.frame(
  term = rep(c("Gaze", "Pupil"), each = 1000),
  draw = c(rnorm(1000, .12, .05), rnorm(1000, -.05, .04))
)
plot_mediation_evidence_atlas(draws, type = "bootstrap")
```

The plotting API does not distinguish bootstrap draws from posterior draws by computation; always label their actual provenance when reporting.

### Compare causal edge assumptions

```{r edges, eval=FALSE}
assumed <- data.frame(from = c("Treatment", "Gaze"),
                      to = c("Gaze", "Choice"))
learned <- data.frame(from = "Treatment", to = "Gaze")
plot_mediation_evidence_atlas(
  type = "dag", theory_edges = assumed, discovered_edges = learned
)
```

**Scientific boundary:** a learned edge is not proof of causation. Retain explicit covariates, measurement timing, randomization assumptions, repeated-measures structure, uncertainty type, and the model that generated the supplied effects. No release is implied by this development vignette.
